---
title: "Explaining poll results by neighborhood characteristics"
author: "Kari Lavikka and Vladimir Dobrodeev"
output: 
  revealjs::revealjs_presentation:
    theme: sky
    transition: slide
    includes:
      in_header: reveal_header.html
---
      
```{r setup, include=FALSE}
# See: https://bookdown.org/yihui/rmarkdown/revealjs.html
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE}
source("../scripts/common.R")

data <- load_data(min_n = 6)

```

## Wind Turbine Poll

Tuulivoiman sosiaalinen hyväksyttävyys Helsingissä  
(Social acceptance of wind power in Helsinki)

https://www.avoindata.fi/data/fi/dataset/helsingin-tuulivoimakysely-2015

### Variables

* Several questions about opinion on turbines
* For example: "Local wind power should be produced in Helsinki"
* Age
* Sex
* Zip code



## Geographical variance in variables

### "Should Helsinki produce local wind power?"

<center>

```{r, message=FALSE, warning=FALSE, out.width="70%", echo=FALSE, results='hide'}
# Load data with supressed output

library(leaflet)
library(sp)
library(rgdal)

postareas <- invisible(spTransform(
  readOGR(file.path("..","raw_data","PKS_postinumeroalueet_2017_shp.shp")),
  CRS("+proj=longlat")
))
postareas$Posno <- as.numeric(as.character(postareas$Posno))
postareas <- postareas[postareas$Posno>=2&postareas$Posno<=990,]

```

```{r, message=FALSE, warning=FALSE, out.width="70%", echo=FALSE}

turbine_zip_data <- read.csv(file.path("..","derived_data","zip_aggregated_turbine.csv"), sep="\t")

names(postareas)[names(postareas)=="Posno"] <- "Zip"

postareas_and_aggregated_data <- merge(postareas, turbine_zip_data, by="Zip")

turbine_poll_questions <- list()

m<-leaflet(postareas_and_aggregated_data) %>% 
  addTiles() %>% 
  fitBounds(24.78516, 60.09772, 25.27679, 60.31403) %>%
  addPolygons(
        weight=1,
        fillColor=~colorNumeric("PiYG",postareas_and_aggregated_data$Local.wind.power.should.be.produced.in.Helsinki)(postareas_and_aggregated_data$Local.wind.power.should.be.produced.in.Helsinki),
        fillOpacity = 0.7,
        layerId = ~Zip
      ) %>%
  addLegend(
    "bottomright",
    title="Average response",
    pal=colorNumeric("PiYG",postareas_and_aggregated_data@data[,8]),
    values=postareas_and_aggregated_data@data[,8],
    opacity=1
  )
m
```
</center>

```{js}
var maps = [];

L.Map.addInitHook(function() {
    var slides = document.querySelector('.slides'),
        zoom = Number(slides.style.zoom);
    maps.push(this);

    // Reveal.js sometimes use the zoom CSS property, but sometimes a scale
    // transform instead. We handle both.
    if (zoom) {
        this._container.style.zoom = 1/zoom;
    } else {
        zoom = Number(slides.style.transform.replace(/.*scale\(([0-9\.]+)\).*/, '$1'));
        this._container.style.transform = 'scale(' + (1 / zoom) + ')';
    }

    this.invalidateSize();
    
    console.log(maps[0]);
    maps[0].on("click", function(e){
      console.log(e);
    });
});
```

## Census data set

Helsinki alueittain  
(Helsinki by regions)

https://www.avoindata.fi/data/fi/dataset/helsinki-alueittain

This data set contains regional demographic data. e.g.

* Support of political parties
* Education
* Employment rate
* Number of Alkos (booze shops)
* And so on



## Geographical variance in variables

### Percentage of families without cars

```{r, results='hide'}

districts <- spTransform(
  readOGR(file.path("..","raw_data","piirialuejako-1995-2016.gpkg"), layer="perus_2016"),
  CRS("+proj=longlat")
)

names(districts)[names(districts)=="PERUS"] <- "District.id"

district_data <- read.csv(file.path("..", "derived_data", "combined_data.csv"), sep="\t", fileEncoding = "utf8", encoding = "utf8")

districts <- merge(districts, district_data, by="District.id")

```

<center>

```{r}
m<-leaflet(districts) %>% 
  addTiles() %>% 
  fitBounds(24.78516, 60.09772, 25.27679, 60.31403) %>%
  addPolygons(
        weight=1,
        fillColor=~colorNumeric("PiYG",districts$Households.without.cars)(districts$Households.without.cars),
        fillOpacity = 0.7,
        layerId = ~District.name
      ) %>%
  addLegend(
    "bottomright",
    title="Per cent of families without cars",
    pal=colorNumeric("PiYG",districts$Households.without.cars),
    values=districts$Households.without.cars,
    opacity=1
  )
m
```

</center>

## Visualizing correlation (1 / 3)

Each point represents a single district

```{r message=FALSE, warning=FALSE, out.width="80%", fig.align="center"}

model_and_plot(data$census$`Swedish speaking`,
               data$turbine$`What do you feel if wind turbines would be built at the outer territorial waters (8-10 km from the coastline) of Helsinki?`,
               list(
                 "Swedish speaking",
                 "Turbines at the outer territorial waters"
               ))

```


## Visualizing correlation (2 / 3)

Each point represents a single district

```{r message=FALSE, warning=FALSE, out.width="80%", fig.align="center"}

model_and_plot(data$census$`Households without cars`,
               data$turbine$`I would like to participate in the construction of wind power as an investor (eg funds, shareholdings).`,
               list(
                 "Households without cars",
                 "Would like to invest in turbines"
               ))

```


## Visualizing correlation (3 / 3)

Each point represents a single district

```{r message=FALSE, warning=FALSE, out.width="80%", fig.align="center"}

model_and_plot(data$census$`Alko per 1000`,
               data$turbine$`Construction of wind turbines would have a positive impact on Helsinki's image.`,
               list(
                 "Number of Alkos per 1000 people",
                 "Turbines have positive impact on the Image of Helsinki"
               ))

```


## Correlation Matrix

```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
plot_correlations()
```

## Conclusions

"Greenpeace" could launch a targeted influencing campaign towards car owners, for example.

* Correlation is not the same as causality
* The method could be applied to other polls as well

Thanks!
