---
title: "Factors that affect poll results"
author: "Kari Lavikka and Vladimir Dobrodeev"
output: 
  revealjs::revealjs_presentation:
    includes:
      in_header: reveal_header.html
  
---

```{r setup, include=FALSE}
# See: https://bookdown.org/yihui/rmarkdown/revealjs.html
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE}
library(readr)

# TODO: Data loading and splitting should be available in a "library"
# So that we can easily include it in 1) pitch, 2) report, 3) Shiny app

combined_data <- read_delim("../derived_data/combined_data.csv", 
                            "\t", escape_double = FALSE, trim_ws = TRUE)

# Subset the data by columns
turbines <- combined_data[, 1:18]
census <- combined_data[, 19:38]

```

## Wind Turbine Poll

Tuulivoiman sosiaalinen hyväksyttävyys Helsingissä  
(Social acceptance of wind power in Helsinki)

https://www.avoindata.fi/data/fi/dataset/helsingin-tuulivoimakysely-2015

### Variables

* Several questions about opinion on turbines
* Age
* Sex
* Zip code



## Geographical variance in variables

### "What do you feel blablaa?"

```{js}


```

<center>

```{r message=FALSE, warning=FALSE, out.width="70%"}
library(leaflet)
library(sp)
library(rgdal)

postareas <- invisible(spTransform(
  readOGR(file.path("..","raw_data","PKS_postinumeroalueet_2017_shp.shp")),
  CRS("+proj=longlat")
))
postareas$Posno <- as.numeric(as.character(postareas$Posno))
postareas <- postareas[postareas$Posno>=2&postareas$Posno<=990,]

turbine_zip_data <- read.csv(file.path("..","derived_data","zip_aggregated_turbine.csv"), sep="\t")

names(postareas)[names(postareas)=="Posno"] <- "Zip"

postareas_and_aggregated_data <- merge(postareas, turbine_zip_data, by="Zip")

turbine_poll_questions <- list()

m<-leaflet(postareas_and_aggregated_data) %>% 
  addTiles() %>% 
  fitBounds(24.78516, 60.09772, 25.27679, 60.31403) %>%
  addPolygons(
        weight=1,
        fillColor=~colorNumeric("PiYG", Helsinki.should.be.an.example.of.reducing.carbon.dioxide.emissions)(Helsinki.should.be.an.example.of.reducing.carbon.dioxide.emissions),
        fillOpacity = 0.7,
        layerId = ~Zip
      )
m
```

</center>

## Census data set

Helsinki alueittain  
(Helsinki by regions)

https://www.avoindata.fi/data/fi/dataset/helsinki-alueittain

This data set contains regional demographic data. e.g.

* Support of political parties
* Education
* Employment rate
* Number of Alkos (booze shops)
* And so on



## Geographical variance in variables

### Percentage of families without cars

TODO: Put a map here




## Area harmonization

* Turbine data has Zip-code areas
* Census data has districts
* These do not line up perfectly
* We used random sampling based on address density within each area


## Visualizing correlation

Each point represents a single district

```{r message=FALSE, warning=FALSE, out.width="70%"}
# TODO: This too should be in a "library"

# Simple linear regression
model_and_plot <- function(x, y, axis_labels) {
  # Read more at: http://r-statistics.co/Linear-Regression.html
  model <- lm(y ~ x)
  plot(x, y,
       pch = 16,
       xlab = axis_labels[[1]],
       ylab = axis_labels[[2]])
  
  abline(model, col = "green")
  
  text(x, y,
       labels = combined_data$`District name`,
       col = "#404040",
       cex = 0.5,
       adj = c(-0.1, 0.5))
  
  title(paste("Correlation:", cor(y, x)))
}

model_and_plot(census$KOK,
               turbines$`I would like to buy local wind power in Helsinki, even if cheaper electricity would be available from elsewhere`,
               list(
                 "Support of Kokoomus",
                 "I would like to by local wind power"))
```


## Conclusions

* "GreenPeace"
* How to target a campaign

Thanks!
