---
title: "Factors that affect poll results"
author: "Kari Lavikka and Vladimir Dobrodeev"
output: 
  revealjs::revealjs_presentation:
    theme: sky
    transition: slide
    includes:
      in_header: reveal_header.html
---
      
```{r setup, include=FALSE}
# See: https://bookdown.org/yihui/rmarkdown/revealjs.html
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE}
source("../scripts/common.R")

data <- load_data(min_n = 6)

```

## Wind Turbine Poll

Tuulivoiman sosiaalinen hyväksyttävyys Helsingissä  
(Social acceptance of wind power in Helsinki)

https://www.avoindata.fi/data/fi/dataset/helsingin-tuulivoimakysely-2015

### Variables

* Several questions about opinion on turbines
* Age
* Sex
* Zip code



## Geographical variance in variables

### "What do you feel blablaa?"

<center>

```{r, message=FALSE, warning=FALSE, out.width="70%", echo=FALSE, results='hide'}
# Load data with supressed output

library(leaflet)
library(sp)
library(rgdal)

postareas <- invisible(spTransform(
  readOGR(file.path("..","raw_data","PKS_postinumeroalueet_2017_shp.shp")),
  CRS("+proj=longlat")
))
postareas$Posno <- as.numeric(as.character(postareas$Posno))
postareas <- postareas[postareas$Posno>=2&postareas$Posno<=990,]

```

```{r, message=FALSE, warning=FALSE, out.width="70%", echo=FALSE}

turbine_zip_data <- read.csv(file.path("..","derived_data","zip_aggregated_turbine.csv"), sep="\t")

names(postareas)[names(postareas)=="Posno"] <- "Zip"

postareas_and_aggregated_data <- merge(postareas, turbine_zip_data, by="Zip")

turbine_poll_questions <- list()

m<-leaflet(postareas_and_aggregated_data) %>% 
  addTiles() %>% 
  fitBounds(24.78516, 60.09772, 25.27679, 60.31403) %>%
  addPolygons(
        weight=1,
        fillColor=~colorNumeric("PiYG",postareas_and_aggregated_data@data[,8])(postareas_and_aggregated_data@data[,8]),
        fillOpacity = 0.7,
        layerId = ~Zip
      ) %>%
  addLegend(
    "bottomright",
    title="Average response",
    pal=colorNumeric("PiYG",postareas_and_aggregated_data@data[,8]),
    values=postareas_and_aggregated_data@data[,8],
    opacity=1
  )
m
```
</center>

```{js}
var maps = [];

L.Map.addInitHook(function() {
    var slides = document.querySelector('.slides'),
        zoom = Number(slides.style.zoom);
    maps.push(this);

    // Reveal.js sometimes use the zoom CSS property, but sometimes a scale
    // transform instead. We handle both.
    if (zoom) {
        this._container.style.zoom = 1/zoom;
    } else {
        zoom = Number(slides.style.transform.replace(/.*scale\(([0-9\.]+)\).*/, '$1'));
        this._container.style.transform = 'scale(' + (1 / zoom) + ')';
    }

    this.invalidateSize();
    
    console.log(maps[0]);
    maps[0].on("click", function(e){
      console.log(e);
    });
});
```

## Census data set

Helsinki alueittain  
(Helsinki by regions)

https://www.avoindata.fi/data/fi/dataset/helsinki-alueittain

This data set contains regional demographic data. e.g.

* Support of political parties
* Education
* Employment rate
* Number of Alkos (booze shops)
* And so on



## Geographical variance in variables

### Percentage of families without cars

TODO: Put a map here




## Area harmonization

* Turbine data has Zip-code areas
* Census data has districts
* These do not line up perfectly
* We used random sampling based on address density within each area


## Visualizing correlation

Each point represents a single district

```{r message=FALSE, warning=FALSE, out.width="80%", fig.align="center"}

model_and_plot(data$census$VASL,
               data$turbine$`Construction of wind turbines would have a positive impact on Helsinki's image.`,
               list(
                 "Support of Vasemmistoliitto",
                 "Turbines would have positive impact on the public image of Helsinki"
               ))

```


## Conclusions

* "GreenPeace"
* How to target a campaign

Thanks!
