---
title: "Explaining Poll Results by Neighborhood Characteristics"
author: "Kari Lavikka and Vladimir Dobrodeev"
date: "23.10.2018"
output:
  pdf_document:
    fig_caption: yes
    toc: true
    number_sections: true
  html_document:
    df_print: paged
references:
- URL: https://www.avoindata.fi/data/fi/dataset/helsinki-alueittain
  id: helsinkiByRegions
  issued:
    year: 2015
  title: Helsinki alueittain
- URL: https://www.avoindata.fi/data/fi/dataset/helsingin-tuulivoimakysely-2015
  id: turbine
  issued:
    year: 2015
  title: Helsingin tuulivoimakysely
- URL: https://openaddresses.io/
  id: openAddresses
  issued:
    year: 2017
  title: OpenAddresses, The free and open global address collection
- URL: https://www.avoindata.fi/data/fi/dataset/helsingin-piirijako
  id: districts
  issued:
    year: 2017
  title: Helsingin piirijako
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
source("../scripts/common.R")
data <- load_data()
```

# Introduction

Various entities conduct polls to find out the public opinion on some matter.
However, it may be difficult to discover the factors that influence the
opinions if no questions related to responders background are asked. Such
background questions could be about common demographic attributes, such as age
and sex, but their explanatory power is limited. Having a poll with an excessive
number of auxiliary questions may be more expensive to conduct, or it may be
less appealing to the potential responders.

Ideally, the poll would only contain the essential questions and the background
information would be obtained from another source. In practice, however, this
is infeasible, as the responders would have to identify themselves so that
results from multiple polls or other data sources could be combined.

We have observed that opinions and demographic characteristics vary between
different geographical regions. It is likely that factors such as accommodation
prices, local culture, or even architecture influence the choice of the
dwelling place.  Based on this observation, we hypothesize that two data sets,
a poll and background data, can be joined by geographical regions. Such
regions may be, for instance, the districts or neighborhoods of a city.

The joined data consists of data points that present the average opinions and
characteristics of a region. By studying how variables correlate among the
regions, some influencing factors may be revealed.

In this project, we study factors that influence public opinion on wind turbine
construction by joining the poll results with a census-dataset. We explain how
the data is joined, visualize certain correlations, and find the most prominent
correlating factors.


# Materials and Methods

## The data sets

Our poll-dataset is about public opinion on wind turbine construction in
Helsinki [@turbine].  The dataset contains individual responses to questions or
claims such as "Wind turbines would influence positively to the public image of
Helsinki.".  However, it contains no background information beside the age, sex
and the *zip code* of the responders.

We join the turbine data with a census dataset [@helsinkiByRegions] that
contains variables such as *mean income*, *families without cars*, *number of
booze shops*, and support for political parties. The data is grouped and
averaged per district.


## Manual data pre-processing

The questions and other variables in the data sets were written in Finnish.  We
translated the relevant parts to English. The census dataset contains over a
hundred variables. We chose a small subset of them for further
processing, which we had considered helpful in explanation of how people responded
or just interesting.


## Geographical area harmonization

The turbine dataset contains *zip codes* of the responders. The census data
set, however, contains data with respect to *districts* of Helsinki. Thus, the geographical areas of
the two data sets are incompatible.  The boundaries of the district and zip
code areas line up partially, but some zip code areas overlap with multiple
districts, and some districts overlap with multiple zip code areas. That has
certain consequences. First, assigning a district for a poll responder is
nontrivial. Second, we can not even be certain about the district when a zip
code area overlaps with two districts.

To assign a district for a responder based on their zip code, we built a set of
conditional probability distributions. The distributions determine at which
probability a responder belongs to a particular district if their zip code is
known.

There are multiple ways to build such distributions. A naive method would be
based on surface areas; how big a zip code area is, and how much it overlaps
with the districts.  However, many regions contain sea or crop fields
and the distribution would not reflect the actual population distribution.

We chose to base the overlap-computation on the distribution of street
addresses [@openAddresses].  Sea or crop fields have no addresses, but
buildings do. We calculate how the coordinates of the street addresses in a
specific zip code area overlap with all known districts [@districts]. Figure
\ref{fig:addresses} shows how the addresses are distributed. Even this method
is not perfect though; some regions have single-family houses dominantly while
others have apartment houses. 

We finally assigned a district for each poll responder by sampling the
conditional probability distributions. Although the district may not be the
true district of the responder, it is still the most likely and spatially close
given the available information.


```{r results="hide"}
#### ALERT ALERT ALERT ###
# TODO: This is copypaste. Instead, create a function
#### ALERT ALERT ALERT ###

library(readr)
library(rgdal)

# Use longlat projection for both the addresses and districts
crs <- CRS("+proj=longlat")

addresses <- read_csv("../raw_data/uusimaa-fi.csv")

# Only retain addresses with a Helsinki postal code
addresses <- addresses[as.numeric(addresses$POSTCODE) < 1000, ]

address_points <- SpatialPoints(with(addresses, cbind(LON, LAT)), crs)
spatial_addresses <- SpatialPointsDataFrame(address_points, addresses)

districts <- readOGR("../raw_data/piirialuejako-1995-2016.gpkg", "perus_2016")

# Fix projection
districts <- spTransform(districts, crs)

# Retain only the addresses that actually reside inside the borders of Helsinki
spatial_addresses <- spatial_addresses[districts, ]
```

```{r dev="png", fig.width=8, fig.height=8, dpi=150, fig.cap="All street addresses within the boundaries of Helsinki. The color indicates the zip code. There are, however, more zip code areas than distinct colors in the scale. The solid lines represent the district boundaries.  \\label{fig:addresses}"}
# Create a palette for plotting
pal <- c(colors(TRUE), colors(TRUE))

# Plot a pretty picture
# First, the addresses as colorful circles
plot(spatial_addresses[districts, ],
     col=pal[as.numeric(addresses$POSTCODE)],
     pch=19,
     cex=0.1
     )
# Second, overlay it with the borders of districts
plot(districts, add=T)
```

## Joining the data sets

After harmonizing the regions, we computed per-district means for the poll
data.  As both of the two data sets now contain district identifiers, we joined
them by the district. As a result, we have a data frame that contains variables
from both the poll results and the census data.

## Visualizing with scatter plot and simple linear regression

To validate our hypotheses and study the correlation between the variables from
the two data sets, we produced scatter plots and fitted a linear model. As the
sample sizes had considerable variation between the district, we adjusted the
point sizes in the plot based on the square root of the sample size. We also
used a weighted linear model to dampen the effect of sampling error that the
small sample sizes cause. Figure \ref{fig:scatterplot} is an example of such a plot.

We also computed weighted correlation coefficients in order to have a single
summary statistic that describes a pair of variables.

```{r fig.height=4.5, fig.cap = "A scatter plot and a regression line. The X and Y axes represent variables from the census and the poll data sets, respectively. \\label{fig:scatterplot}"}

model_and_plot(data$census$VASL,
               data$turbine$`Construction of wind turbines would have a positive impact on Helsinki's image.`,
               axis_labels = list(
                 "Support of Vasemmistoliitto",
                 "Turbines would have positive impact..."
               ))
```


## Correlation matrix

Instead of visualizing all possible variable combinations with scatter plots,
we computed a weighted correlation matrix. We also computed significance levels
for the correlations and visualized the results using a correlation plot.  The
correlation matrix is further discussed in Chapter \ref{chap:results}.


## Variable selection with regularized multiple linear regression

Correlation, being similar to covariance, describes how two random variables
vary together. However, because only two variables are involved, it can not
reveal more complex phenomenon, such as how certain factors jointly influence
an opinion.

To get some insights into the joint influence, we normalized the variables and
used multiple linear regression to reveal the factors that have either clearly
positive or negative correlation with the opinions. The model was regularized
using the lasso method, and the $\lambda$-value was determined by
cross-validation.  Figure \ref{fig:multipleRegression} shows an example of
variable selection on normalized data.


```{r fig.height=3.5, fig.cap="Coefficients of a regularized multiple linear regression.\\label{fig:multipleRegression}"}

explain_variable(data$turbine$`What do you feel if wind turbines would be built at the outer territorial waters (8-10 km from the coastline) of Helsinki?`, data)
title("What do you feel if wind turbines would be built at the outer territorial waters of Helsinki?", cex.main = 0.7)

```

# Results
\label{chap:results}

The results are best summarized by a picture. Figure
\ref{fig:correlationMatrix} displays a correlation matrix. The color encodes the
correlation, and the size of a circle encodes the magnitude of the correlation.
Circles are only shown for variable pairs that have significant correlation,
i.e. $\text{p-value} < 0.05$. 

The factors *secondary education*, *households without cars*, *support of the
Green Party* (VIHR), and *support of Vasemmistoliitto* (VASL) have clearly
positive correlation with all questions related to turbine construction. In
contrast, *mean income*, *support of RKP*, and especially *support of Kokoomus*
(KOK) have a strong negative correlation with the poll questions.

The correlation matrix displays only a slight variation between the questions.
The regularized multiple linear regression, on the other hand, gives highly
variable results. Some of them, such as the question shown in Figure
\ref{fig:multipleRegression}, mostly align with the correlation matrix.  Others
produce very different results.

```{r fig.cap="The correlation matrix. Rows represent poll questions, columns represent census variables. \\label{fig:correlationMatrix}"}
plot_correlations(data)
```

# Discussion

The results suggest that placement on the political axis strongly correlates
with the opinion on construction of wind turbines.  The Green Party and
Vasemmistoliitto represent left-wing politics, and according to the results,
tend to support wind turbine construction.  Conversely, Kokoomus and RKP
represent right-wing politics and tend to oppose turbine construction. Other
factors are of less importance but may warrant further investigation.

We have demonstrated that our method is able to reveal factors that correlate
with the opinions on wind turbines. However, we are playing with population
averages of only 34 districts, and weaker signals will get lost in the process.
Also, we have not proven our method formally.

The joined data along with the analyses presented here are available as an
interactive web app: https://tuner.shinyapps.io/turbine-analysis/

The source code for the web app and this report is available on GitHub:
https://github.com/tuner/turbine-analysis


# Authors' contributions

Kari Lavikka came up with the project idea, harmonized the regions, analyzed
correlations and regressions, and drafted the report.  Vladimir Dobrodeev contributed to
the project idea and design, implemented geographical visualizations,
implemented the Shiny web-application, and revised the report.  Both
contributed equally to the elevator-pitch presentation.


# References
